---
title: "Question 2"
author: "Eva Burguete-Innocente & Elliot Gavrin"
date: "`r Sys.Date()`"
output: html_document
---

Question 1a: How does heritability change as you move up the taxonomic rank?
1b: How do the different variables (domestication, thermy, vertebrates etc.) affect heritability?
Hypothesis 1a: Heritability will increase/decrease as you move up the taxonomic rank.
1b: Heritability will change based on the different levels of vertebrate, thermy, domestication, and lab/field conditions. 

Loading necessary packages:

```{r}
library(ggplot2)
library(dplyr)
library(stringr)
library(lmtest)
library(lme4)
library(tidyr)
library(RColorBrewer)
```

Reading in the data and checking it:

```{r}
data<-read.csv("Data/Compiled2.csv")
head(data)
str(data)
```

Using regular expressions to filter out mistakes in the data:

```{r}
data%>%
  filter(str_detect(Genus,"^T"))

#remove Tachymarptis\xa0 and replace with Tachymarptis

data$Genus<-gsub("Tachymarptis\xa0","Tachymarptis",data$Genus)

```

Plotting the effects of other variables on heritability. 

```{r}
ggplot(data,aes(L.F,Zr))+geom_boxplot()

ggplot(data,aes(Thermy,Zr))+geom_boxplot()

ggplot(data,aes(Vert,Zr))+geom_boxplot()

ggplot(data,aes(D.W,Zr))+geom_boxplot() 
```

Plotting how Zr differs across taxa and within taxonomic ranks.

```{r}
ggplot(data,aes(Phylum,Zr,fill=Phylum))+geom_boxplot() 

ggplot(data,aes(Class,Zr,fill=Class))+geom_boxplot()

ggplot(data,aes(Order,Zr,fill=Order))+geom_boxplot()+theme(legend.position="none")

ggplot(data,aes(Family,Zr,fill=Family))+geom_boxplot()+theme(legend.position="none")

ggplot(data,aes(Genus,Zr,fill=Genus))+geom_boxplot()+theme(legend.position="none")

ggplot(data,aes(species.epithet,Zr,fill=species.epithet))+geom_boxplot()+theme(legend.position="none")
```

Creating data for plotting and plotting the average heritability of each group within each taxonomic rank. 

```{r}

pivotdat<-data%>%
  select(Phylum,Class,Order,Family,Genus,species.epithet,Zr)%>%
  group_by(Phylum,Class,Order,Family,Genus,species.epithet)%>%
  mutate(meanZr=mean(Zr))

longdat<-pivot_longer(pivotdat,cols=c("Phylum","Class","Order","Family","Genus","species.epithet"),names_to="Rank",values_to="Name")

longdatplot<-longdat%>%
  group_by(Rank,Name)%>%
  mutate(meanZr=mean(Zr))%>%
  select(Rank,Name,meanZr)

ggplot(longdatplot,aes(Rank,meanZr,fill=Rank))+geom_boxplot()+theme(legend.position="none")

phylum<-longdatplot%>%
  filter(Rank=="Phylum")%>%
  summarise(mean(meanZr))

class<-longdatplot%>%
  filter(Rank=="Class")%>%
  summarise(mean(meanZr))

order<-longdatplot%>%
  filter(Rank=="Order")%>%
  summarise(mean(meanZr))

family<-longdatplot%>%
  filter(Rank=="Family")%>%
  summarise(mean(meanZr))

genus<-longdatplot%>%
  filter(Rank=="Genus")%>%
  summarise(mean(meanZr))

species<-longdatplot%>%
  filter(Rank=="species.epithet")%>%
  summarise(mean(meanZr))

plotdat2<-rbind(phylum,class,order,family,genus,species)
plotdat2<-plotdat2%>%
  mutate(avg=`mean(meanZr)`)%>%
  select(Rank,Name,avg)

rankplot<-ggplot(plotdat2,aes(Rank,avg,fill=Rank))+geom_boxplot()+theme_bw()+ylab("Average Heritability")+xlab("Taxonomic Rank")+scale_x_discrete(limits = c("Phylum", "Class","Order","Family","Genus","species.epithet"),labels= c("Phylum", "Class","Order","Family","Genus","Species"))+scale_fill_brewer(palette="YlOrRd")+theme(legend.position="none")
rankplot
```

Creating linear mixed effects models including multiple combinations of variables and running a likelihood to see which is the best. 

```{r}
MixedModel <- lmer(Zr ~ Vert*Thermy*D.W*L.F + (1|Study.ID) + (1|Family/Genus/species.epithet), data = data)

summary(MixedModel)

MixedModel1 <- lmer(Zr ~ Vert*Thermy*D.W + (1|Study.ID)+(1|Family/Genus/species.epithet), data = data) 

MixedModel1a <- lmer(Zr ~ Vert*Thermy*L.F + (1|Study.ID)+(1|Family/Genus/species.epithet), data = data)

MixedModel1b <- lmer(Zr ~ Vert*D.W*L.F + (1|Study.ID) +(1|Family/Genus/species.epithet),data = data)

MixedModel1c <- lmer(Zr ~ Vert*Thermy*D.W + (1|Study.ID)+ (1|Family/Genus/species.epithet),data = data) #Best Model according to lrtest

MixedModel1d <- lmer(Zr ~ Thermy*D.W*L.F + (1|Study.ID)+(1|Family/Genus/species.epithet), data = data)

MixedModel2 <- lmer(Zr ~ Vert*Thermy + (1|Study.ID)+ (1|Family/Genus/species.epithet),data = data)

MixedModel2a <- lmer(Zr ~ Vert*D.W + (1|Study.ID)+ (1|Family/Genus/species.epithet),data = data)

MixedModel2b <- lmer(Zr ~ Vert*L.F + (1|Study.ID)+ (1|Family/Genus/species.epithet),data = data)

MixedModel2c <- lmer(Zr ~ D.W*Thermy + (1|Study.ID)+ (1|Family/Genus/species.epithet),data = data)

MixedModel2d <- lmer(Zr ~ L.F*Thermy + (1|Study.ID)+(1|Family/Genus/species.epithet),data = data)

MixedModel2e <- lmer(Zr ~ L.F*D.W + (1|Study.ID)+ (1|Family/Genus/species.epithet),data = data) #Best Model according to lrtest

MixedModel3 <- lmer(Zr ~ Vert + (1|Study.ID)+(1|Family/Genus/species.epithet), data = data)

MixedModel3a <- lmer(Zr ~ Thermy + (1|Study.ID)+(1|Family/Genus/species.epithet), data = data) #Best Model according to lrtest

MixedModel3b <- lmer(Zr ~ D.W + (1|Study.ID)+(1|Family/Genus/species.epithet), data = data)


MixedModel3c <- lmer(Zr ~ L.F + (1|Study.ID)+(1|Family/Genus/species.epithet), data = data)

lrtest(MixedModel, MixedModel1, MixedModel1a, MixedModel1b,MixedModel1c,MixedModel1d,MixedModel2,MixedModel2a,MixedModel2b,MixedModel2c,MixedModel2d,MixedModel2e, MixedModel3,MixedModel3a,MixedModel3b,MixedModel3c) 
```

Using AIC and BIC to evaluate how well the models perform. 

```{r}
AIC(MixedModel, MixedModel1, MixedModel1a, MixedModel1b,MixedModel1c,MixedModel1d,MixedModel2,MixedModel2a,MixedModel2b,MixedModel2c,MixedModel2d,MixedModel2e, MixedModel3,MixedModel3a,MixedModel3b,MixedModel3c)

BIC(MixedModel, MixedModel1, MixedModel1a, MixedModel1b,MixedModel1c,MixedModel1d,MixedModel2,MixedModel2a,MixedModel2b,MixedModel2c,MixedModel2d,MixedModel2e, MixedModel3,MixedModel3a,MixedModel3b,MixedModel3c)
```

Plotting lmer models and plotting the effects of the fixed effects on heritability. 

```{r}
library(sjPlot) #for plotting lmer models
tab_model(MixedModel1c,
                  show.re.var= TRUE,
                  pred.labels =c("(Intercept)", "Vertebrate", "Endotherm", "Semi-domesticated", "Wild", "Vertebrate × Wild", "Endotherm × Semi-domesticated"),
                  dv.labels= "Effects of Fixed Effects on Heritability")

p <-sjPlot::plot_model(MixedModel1c,axis.labels=c("Endotherm × Semi-domesticated", "Vertebrate × Wild", "Wild", "Semi-domesticated", "Endotherm", "Vertebrate", "Endotherm × Semi-domesticated"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effects of Fixed Effects on Heritability")
p+theme_sjplot()
```

